OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-06-15 paper/paper.Rmd
Summarize results in abstract

Assigning amplicon sequences to Operational Taxonomic Units (OTUs) is an
important step in characterizing the composition of microbial communities across
large datasets.
OptiClust, a _de novo_ OTU clustering method in the mothur program, has been
shown to produce higher quality OTU assignments than other methods and at
comparable or faster speeds [@westcott_opticlust_2017;
@schloss_introducing_2009].
A notable difference between _de novo_ clustering and database-dependent methods
is that OTU assignments clustered with _de novo_ methods are not stable when new
sequences are added to a dataset [@westcott_novo_2015].
However, in some cases one may wish to incorporate new samples into a previously
clustered dataset without performing clustering again on all sequences, such as
when deploying a machine learning model where OTUs are features
[@topcuoglu_effective_2019].
To provide an efficient and robust method to fit amplicon sequence data to
existing OTUs, we developed the OptiFit algorithm as a new component of the
mothur program.

**TODO: summarize results & conclusion**

>>>>

Assigning amplicon sequences to Operational Taxonomic Units (OTUs) is an
important step in characterizing the composition of microbial communities across
large datasets.
OptiClust, a _de novo_ OTU clustering method in the mothur program, has been
shown to produce higher quality OTU assignments than other methods and at
comparable or faster speeds [@westcott_opticlust_2017;
@schloss_introducing_2009].
A notable difference between _de novo_ clustering and database-dependent methods
is that OTU assignments clustered with _de novo_ methods are not stable when new
sequences are added to a dataset [@westcott_novo_2015].
However, in some cases one may wish to incorporate new samples into a previously
clustered dataset without performing clustering again on all sequences, such as
when deploying a machine learning model where OTUs are features
[@topcuoglu_effective_2019].
To provide an efficient and robust method to fit amplicon sequence data to
existing OTUs, we developed the OptiFit algorithm as a new component of the
mothur program.
We tested OptiFit using four microbiome datasets with two different strategies;
by fitting to an external reference database, or by splitting the dataset into a 
reference and query set and fitting the query sequences to the reference set.
OptiFit produces OTUs of similar quality as OptiClust and at faster speeds when
using the split dataset strategy, while the OTU quality and speed depends on the
database chosen when using the external database strategy.
OptiFit provides a suitable option for users who require stable reference-based
OTU assignments at the same quality afforded by _de novo_ clustering methods.
